---
title: 'DSO 545: HW 1'
author: "Bradley Rava, Patrick Vossler, Simeng Shao"
date: "1/27/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)

```

Load the data:
```{r}

baggage = read.csv(here("HW1", "Baggage.csv"), header=T,stringsAsFactors = F)
indus_med = read.csv(here("HW1","IndustryMedians.csv"),header=T)
head(baggage)
```

Process data:
```{r}
baggage$Date = as.Date(paste0("02/",baggage$Date),"%d/%m/%Y")
baggage$Month = factor(baggage$Month,labels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
baggage$Airline = as.character(baggage$Airline)
```

## 1. Explore baggage complaints over time: create 3 time series plots for the variable *Baggage* by Date for each of the airlines separately.
```{r}
airlines = unique(baggage$Airline)
for(i in 1:length(airlines)){
    airline = airlines[i]
    data = baggage[baggage$Airline == airline,]
    res = aggregate(data["Baggage"], by=list(Month = data$Month), sum)
    plot(x=as.integer(res$Month),y=res$Baggage,type="o",xaxt="n",xlab="Month", ylab="Number of Baggage Complaints")
    axis(1,at = seq(1,12),labels = levels(res$Month))
    title(paste(airline,"Baggage Complaints (2004-2010)"))
    
}
```

## 2. Briefly describe what patterns you see in the plots
In some of the plots we see a cyclical pattern with the number of baggage complaints increasing during the winter holiday travel season (November-January). There is often another spike in baggage complaints in the summer likely when families are going on summer vacations.

* American Eagle
    * We see that the cyclical yearly trend described above holds for American Eagle. Furthermore we see that there is an increase in the total number of complaints in 2006-2008 and then the number of complaints drops back down from 2009 onward.
* Hawaiian Airlines
    * Compared to American Eagle, Hawaiian Airlines has a smaller number of complaints each month. This is expected because Hawaiian Airlines is a smaller airline compared to American Eagle. Whereas American Eagle had a spike in baggage complaints during the winter holiday travel season, Hawaiian Airlines seems to have spikes in baggage complaints during the Spring and Summer. This perhaps could be because they see an influx of passengers wishing to travel to Hawaii during the Spring and Summer months.
    * The most concerning trend for Hawaiian Airlines is the trend of larger spikes in each of the successive years, culminating with a large spike in baggage complaints during the 2010 holiday season.
* United Airlines
    * Unsurprisingly United Airlines has a larger number of baggage complaints overall which can be explained by its much larger size compared to the other two companies.
    * Like American Eagle we see that United Airlines also experiences a surge in baggage claims during the holiday season. Additionally, it is interesting that both American Eagle and United Airlines have a spike in baggage complaints during 2006. Perhaps there was some external event that caused this for both airlines?
    * Since both American Eagle and United Airlines provide a variety of flights to domestic destinations it is not surprising to see that they have similar baggage complaint patterns in the summer and winter months.

## 3.
```{r}
airlines = unique(baggage$Airline)
for(i in 1:length(airlines)){
    airline = airlines[i]
    data = baggage[baggage$Airline == airline,]
    res = aggregate(data["Baggage"], by=list(Month = data$Month, Year = data$Year), sum)
    years = unique(data$Year)
    plot_dat = res[res$Year == years[1],]
    
    #bottom,left,top,right margin
    par(mar=c(7.1, 4.1, 3.1, 8.9), xpd=TRUE)

    plot(x=as.integer(plot_dat$Month),y=plot_dat$Baggage,type="o",xaxt="n",xlab="", ylab="Number of Baggage Complaints",lty=1, col=1, pch = 1,ylim = c(min(res$Baggage),max(res$Baggage)))
    axis(1,at = seq(1,12),labels = levels(res$Month))
    title(paste(airline,"Baggage Complaints"))   
    for(j in 1:length(years)){
        plot_dat = res[res$Year == years[j],]
        lines(x=as.integer(plot_dat$Month),y=plot_dat$Baggage,type="o",lty=j, col=j,pch=j)
        
    }
    legend("topright", inset=c(-0.2,0), legend=years, pch=1:length(years),lty=1:length(years),col=1:length(years), title="Years")
    
    
}
```

## 4. Describe the patterns in the plot

## 5. Plot all three airline Baggage data by Date on one graph.
```{r}
# Maybe do this on the log scale?
airlines = unique(baggage$Airline)
airline = airlines[1]

total_aggregated = aggregate(baggage["Baggage"], by=list(Date = baggage$Date,Airline=baggage$Airline), sum)

res = total_aggregated[baggage$Airline == airline,]



par(mar=c(7.1, 4.1, 3.1, 8.9), xpd=TRUE)

plot(x=res$Date,y=res$Baggage,type="o",xaxt="n",xlab="Month", ylab="Number of Baggage Complaints",lty=1, col=1, pch = 1,ylim= c(0,max(total_aggregated$Baggage)))

axis(1, res$Date, format(res$Date, "%b %d"))
title("Baggage Complaints for all 3 Airlines (2004-2010)")
for(i in 2:length(airlines)){
    airline = airlines[i]
    res = total_aggregated[baggage$Airline == airline,]
    lines(x=res$Date,y=res$Baggage,type="o",lty=i, col=i,pch=i)
}
legend("topright", inset=c(-0.375,0), legend=airlines, pch=1:length(airlines),lty=1:length(airlines),col=1:length(airlines), title="Airlines")

```

## 6. Based on the graph in question 5., do some airlines have better baggage handling practices?
According to the plot, Hawaiian line seems to have much smaller baggage complaints throughout 2004-2010, which has been below 50000. Other two lines, however, are above 50000.

## 7.	Based on the graph in question 5., which airline has the best record? The worst?
Based on the graph, Hawaiian has the best record, United has the worst record.


## 8.	Based on the graph in question 5., are complaints getting better or worse over time?

## 9.	Are the conclusions, you have drawn based on the graphs of the raw data you created, accurate? Are there any potential factors that may distort your conclusions and should be taken into consideration?

## 12. Create **Baggage %** KPI
```{r}
baggage$Baggage_perc = baggage$Baggage / baggage$Enplaned * 100
```


## 17. Superimpose time series plots of monthly averages of **Baggage %** by time for the three airlines
Plot the mean of each month for all of the years


## 19. Create a timeplot of **Baggage %**, add average line for **Baggage %** and a trendline of monthly average **Baggage %** for each airline.
Create monthly averages over total data and plot lm


# Case 2: CEO Compensation

Load CEO compensation data. Using a csv form of the data because I was getting an error when trying to read the tab separated file
```{r}
ceo_comp = read.csv(here("HW1","CEOcompensation.csv"),header=T)
```


## 1. What is the number of female CEOs?
```{r}
nrow(ceo_comp[ceo_comp$Gender=="F",])
```


## 2. What is the age of the youngest CEO?
```{r}
min(ceo_comp$Age)
```

## 3. What is the age of the oldest CEO?
```{r}
max(ceo_comp$Age)
```

## 4. What is the average age of a CEO?
```{r}
round(mean(ceo_comp$Age),2)
```
## 5. What is the total CEO 2008 salary?
```{r}
round(sum(ceo_comp$X2008.Salary),2)
```

## 6. How many CEOs have joined a company as a CEO?
So they weren't a founder and they weren't at the company before becoming the CEO
```{r}
# sum(ceo_comp$Founder == "No" & ceo_comp$Years.as.company.CEO == ceo_comp$Years.with.company)
sum(ceo_comp$Years.as.company.CEO == ceo_comp$Years.with.company)
```

## 7. What is the average amount of time a CEO worked for a company before becoming a CEO?
```{r}
worked_before = ceo_comp[ceo_comp$Years.with.company > ceo_comp$Years.as.company.CEO,]
round(mean(worked_before$Years.with.company - worked_before$Years.as.company.CEO),2)
```


## 8. Which industry in the data set has the largest number of CEO's
```{r}
which.max(table(ceo_comp$Industry))
```
Oil & Gas Operations has the largest number of CEO's in the data set



